---
title: "App Starter"
output:
  flexdashboard::flex_dashboard:
    theme:
      bootswatch: flatly
    orientation: rows
    social: menu
    source_code: embed
runtime: shiny
---

```{r, echo = FALSE, warning=FALSE, message=FALSE}
library(flexdashboard)

library(tidyverse)
library(forecast)
library(plotly)

# historical records from multiple stores
fsales <- "https://raw.githubusercontent.com/multidis/hult-inter-bus-reports-r/main/forecasting/sales_weekly.csv"
sales <- read_csv(fsales)

# latest (current) week
nweek_now <- max(sales$Week)
```


Static chart {data-icon="fa-line-chart"}
=========================================

Sidebar {.sidebar}
-------------------------------------

```{r echo = FALSE}
selectInput("storenum0", label = "Select store: ",
            choices = 1:max(sales$Store), selected = 1)

sliderInput("horizon0", label = "Forecast horizon (weeks): ",
            min = 2, max = 52, value = 8, step = 1)
```


Main content
-----------------------------------------

Please note that all transformations responding to inputs must be included in **reactive** blocks, typically called `render...` (e.g. `renderPlot`, `renderTable` etc.).

```{r echo = FALSE}
renderPlot({
  horiz <- input$horizon0
  
  sales_ts <- ts(sales %>% filter(Store == input$storenum0) %>%
                   select(Weekly_Sales), frequency = 52)
  arima_model <- auto.arima(sales_ts, seasonal.test = "seas")
  arima_pred <- forecast(arima_model, h = horiz)
  
  sales_pred <- data.frame(Sales_Predicted = as.numeric(arima_pred$mean),
                           Week = (nweek_now + 1):(nweek_now + horiz))
  sales_pred %>% ggplot(aes(Week, Sales_Predicted)) +
    geom_line() + labs(x = "Weeks ahead", y = "Weekly Sales Forecast")
})
```


Interactive chart {data-icon="fa-mouse-pointer"}
==================================================

Please note that all transformations responding to inputs must be included in **reactive** blocks, typically called `render...` (e.g. `renderPlot`, `renderTable` etc.).

### Selectors

```{r echo = FALSE}
selectInput("storenum", label = "Select store: ",
            choices = 1:max(sales$Store), selected = 1)

sliderInput("horizon", label = "Forecast horizon (weeks): ",
            min = 2, max = 52, value = 8, step = 1)
```


### Chart header
    
```{r echo = FALSE}
output$plotly <- renderPlotly({
  horiz <- input$horizon
  
  sales_ts <- ts(sales %>% filter(Store == input$storenum) %>%
                   select(Weekly_Sales), frequency = 52)
  arima_model <- auto.arima(sales_ts, seasonal.test = "seas")
  arima_pred <- forecast(arima_model, h = horiz)
  
  sales_pred <- data.frame(Sales_Predicted = as.numeric(arima_pred$mean),
                           Week = (nweek_now + 1):(nweek_now + horiz))
  
  p <- sales_pred %>% ggplot(aes(Week, Sales_Predicted)) +
    geom_line() + labs(x = "Weeks ahead", y = "Weekly Sales Forecast")
  
  cd <- session$clientData
  ggplotly(p, height = cd$output_plotly_height, width = cd$output_plotly_width)
})

plotlyOutput("plotly")
```
