---
title: "Exploring web applications"
author: "Int Rep with R"
date: "2023-07-09"
output: html_document
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
```

You can always modify this text as you like.

## Inputs and Outputs

You can embed Shiny inputs and outputs in your document. Outputs are automatically updated whenever inputs change.  This demonstrates how a standard R plot can be made interactive by wrapping it in the Shiny `renderPlot` function. The `selectInput` and `sliderInput` functions create the input widgets used to drive the plot.

```{r eruptions, echo=FALSE}
inputPanel(
  selectInput("n_breaks", label = "Number of bins:",
              choices = c(10, 20, 35, 50), selected = 20),
  selectInput("numrows", label = "Number of rows in the table:",
              choices = c(15, 25, 50), selected = 25),
  checkboxInput("cbox", label = "Select here:", value = TRUE)
)
```

Checkbox state

```{r, echo=FALSE}
renderPrint(input$cbox)
```

Now the table

```{r, echo=FALSE}
conditionalPanel(
  condition = "input.cbox",
  renderDataTable(faithful[1:input$numrows, ])
)
```

Back to the histogram

```{r, echo=FALSE}
renderPlot({
  hist(faithful$eruptions, probability = TRUE, breaks = as.numeric(input$n_breaks),
       xlab = "Duration (minutes)", main = "Geyser eruption duration")
  
  dens <- density(faithful$eruptions, adjust = 1.2)
  lines(dens, col = "blue")
})
```

Some more text here.

