---
title: "Storytelling in Data Analytics"
subtitle: "Building Interactive Business Reports with R (DAT-3535)"
author: "Beau Giannini and Pavel Paramonov"
date: "Hult International Business School"
output:
  tufte::tufte_html: default
---

```{r setup, include=FALSE}
library(tufte)
# invalidate cache when the tufte version changes
knitr::opts_chunk$set(tidy = FALSE, cache.extra = packageVersion('tufte'))
options(htmltools.dir.version = FALSE)

library(tidyverse)
library(ggplot2)
library(ggtext)

source("theme_swd.R")
```

`r newthought('An example analysis')` provided in this handout is designed to illustrate the storytelling aspect of business reporting. Your task is to take it as a starting point and produce a report with the best possible layout, appeal, and interactivity that are possible to achieve within a static HTML output format.

You may find the following resources useful^[You are not in any way restricted to use these tools only - explore the R ecosystem!] to the chart types presented in this example:

- Interactive `ggplot2` figures using [plotly](https://plotly.com/ggplot2/).
- [htmlwidgets](https://www.htmlwidgets.org/) for R.
- [flexdashboard](https://pkgs.rstudio.com/flexdashboard/) system for R integrated with RStudio tools.
- Links and materials posted on MyCourses in the respective modules.

You should strive to produce appropriately designed, clear, and uncluttered visualizations combined with narratives that explain an important aspect of the analysis and engages the audience.


## Airline On-Time Arrival Performance

On-time and delayed flight data by airline for the San Francisco International Airport (SFO).

- Time period Jan 1 through Dec 31, 2019.^[Your data description should be specific enough to know the time periods, geographical characteristics, population specifics and/or any other aspects essential for the dataset.]
- All major airlines arriving to SFO are included.
- A flight is considered delayed if the arrival time lags the schedule by 15 minutes or more.
- Counts of delayed flights are provided for various delay causes.

*Data source:* US Department of Transportation, Bureau of Transportation Statistics. [https://www.transtats.bts.gov/OT_Delay/OT_DelayCause1.asp](https://www.transtats.bts.gov/OT_Delay/OT_DelayCause1.asp). (Retrieved Feb 24, 2020)^[A proper citation of the dataset origin along with the web link is always helpful and adds credibility.].


## Flight delays by airline

```{r, echo=FALSE, warnings=FALSE, message=FALSE, fig.width=6, fig.height=4}
load(file = "handout_prep_out.rda")

flight_delay_comp %>%
   ggplot(aes(x=reorder(name, delayedpct), y=delayedpct/100, fill = highlight)) +
   geom_col(width = 0.8) + coord_flip(clip = "off") +
   scale_x_discrete() +
   scale_y_continuous(position = "right", limits = c(0, 0.5), labels = scales::percent) +
   scale_fill_manual(values = c(GRAY9, RED3), guide = "none") +
   labs(
      caption = "SFO arrivals, 2019.\nData source: US Department of Transportation.",
      y = "% flights delayed over 15 minutes",
      title = "**Frontier Airlines** has the largest delayed flights fraction"
   ) +
   theme_swd() + theme(
      axis.title.y = element_blank(),
      plot.title = element_markdown(lineheight = 1.2),
      plot.subtitle = element_markdown(lineheight = 1.5, linewidth = 10),
      axis.text.x = element_text(size = 9, color = GRAY5),
      axis.text.y = element_text(color = GRAY4, size = 10),
      axis.title.x = element_text(color = GRAY5, size = 10, hjust = 0.05),  
      axis.line.y = element_blank(),
      axis.ticks.y = element_blank(),
      strip.placement = "outside",
      strip.background = element_rect(fill = NA, color = GRAY9)
   )
```


## Flight numbers by airline

```{r, echo=FALSE, fig.width=6, fig.height=4}
flight_number_comp %>%
   ggplot(aes(x=reorder(name, total), y=total/1000, fill = highlight)) +
   geom_col(width = 0.8) + coord_flip(clip = "off") +
   scale_x_discrete() +
   scale_y_continuous(position = "right", limits = c(0, 60)) +
   scale_fill_manual(values = c(GRAY9, RED3), guide = "none") +
   labs(
      caption = "SFO arrivals, 2019.\nData source: US Department of Transportation.",
      y = "Total flights (thousands)",
      title = "The number of Frontier Airlines flights is among the fewest in SFO"
   ) +
   theme_swd() + theme(
      axis.title.y = element_blank(),
      plot.title = element_markdown(lineheight = 1.2),
      plot.subtitle = element_markdown(lineheight = 1.5, linewidth = 10),
      axis.text.x = element_text(size = 9, color = GRAY5),
      axis.text.y = element_text(color = GRAY4, size = 10),
      axis.title.x = element_text(color = GRAY5, size = 10, hjust = 0.05),  
      axis.line.y = element_blank(),
      axis.ticks.y = element_blank(),
      strip.placement = "outside",
      strip.background = element_rect(fill = NA, color = GRAY9)
   )
```

While Frontier Airlines have the largest **fraction** of delayed flights, this is not related to a large number of flights^[It may help to choose a visualization that addresses some potential criticisms of your initial conclusion. For instance, the audience may claim that frontier has more delays because it operates a much larger number of flights. This chart clearly shows that is not the case.]. In fact, most other carriers operate a significantly larger number of flights arriving to SFO.


## Flight delay causes

```{r, echo=FALSE, fig.width=8, fig.height=5.5}
ggplot(df, aes(x = factor(Airline, levels = c("Frontier", "Others")), y = value,
                            fill = factor(cause, levels = c("Carrier", "Aircraft", "Cancelled",
                                                            "Aviation", "Other")))) + 
   geom_col(position = "stack", color = "white", width = .6) + 
   scale_fill_manual(name="Cause", values = c(BLUE2,ORANGE2,RED3,GRAY9,GRAY2)) + 
   scale_y_continuous(expand = c(0,0), labels = scales::percent) +
   coord_cartesian(clip = "off") + 
   labs(title = "Causes of flight delays: Frontier vs. Others") + 
   theme_minimal() + theme(#panel.grid.major = element_blank(),
      #panel.grid.minor = element_blank(),
      #axis.line.y = element_blank(),
      axis.line.x = element_line(color = GRAY9),
      axis.title = element_blank(),
      #axis.ticks = element_blank(),
      #axis.text.y = element_blank(),
      axis.text.x = element_text(size = 12),
      plot.margin = unit(c(1,1,1,4), "cm"),
      plot.title = element_text(size = 18, color = GRAY2, margin = margin(0,0,.5,0,"cm")),
      plot.caption = element_text(color = GRAY8, hjust = 0, margin = margin(.3,0,0,0,"cm"))
   )
```

Similar factors cause flight delays for Frontier and for other airlines.


## Source code

You can find all files needed to produce the current report in the [Github class repository](https://github.com/multidis/hult-inter-bus-reports-r/tree/main/reporting). Please make sure to save all of the files to your working directory.
